---
title: "Coverage analysis of cttso data using TST500C_manifest.bed"
author: "Sehrish Kanwal"
date: "`r format(Sys.time(), '%a %Y-%b-%d')`"
output:
  html_document:
    keep_md: yes
    css: ../style.css
    theme: readable
    toc: true
    toc_float: true
    code_folding: hide
  rmdformats::material:
    highlight: kate
params:
  mosdepth_output: '/Users/kanwals/UMCCR/data/projects/cttso-qc/stitched_bams_mosdepth_output_tst500c'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
### Required packages.
library(here)
library(dplyr)
library(tidyverse)
library(stringr)
library(DT)
library(jtools)
```

***

## Introduction

In this analysis, we are looking at the coverage of panel target regions across our tso cohort.

## Data

The TSO500 alignment files used for calculating coverage exist in GDS. The script used to run [mosdepth tool](https://github.com/brentp/mosdepth) on these bams for ~311 samples can be found [here](./scripts/run_cttso_bams_through_mosdepth.sh)

mosdepth can output the mean per-region given a BED file of regions. The result is summarized below.


```{r raw_cov_results, message=FALSE, warning=FALSE}
# read in bed file with regions of interest
bed_coverage <- as.data.frame(read.table("/Users/kanwals/UMCCR/data/projects/cttso-qc/TST500C_manifest.bed", header = FALSE, sep="\t", stringsAsFactors=FALSE))
colnames(bed_coverage) <- c("chr", "region_start", "region_end", "name")

# list files in the input directory - made using mosdepth
list_of_files <- list.files(path = params$mosdepth_output,
                            recursive = TRUE,
                            pattern = "\\.bed.gz$",
                            full.names = TRUE)

for (i in 1:length(list_of_files)){
  # read files and prepare data format for display
  file <- read_tsv(list_of_files[i],col_names = c("chr", "region_start", "region_end", "name", "mean_coverage" ),col_types = "cn") %>%
      mutate(sample_id = basename(list_of_files[i])) %>%
      mutate(sample_id = gsub("*.regions.bed.gz", "", sample_id)) 
  # find sample id and rename mean_coverage column using it
  sample_name <- unique(file$sample_id)
  sample_coverage <- file %>%
    dplyr::rename(!!sample_name := mean_coverage) %>%
    dplyr::select(one_of(sample_name))
  # add sample_coverage to original bed file
  bed_coverage <- cbind(bed_coverage, sample_coverage)
}

mean_cov_across_samples = round(rowMeans(bed_coverage[,c(5:ncol(bed_coverage))]), 2)
bed_coverage <- add_column(bed_coverage, mean_cov_across_samples, .after = 4)

```

## Coverage check

We are interested to check what regions consistently have less than X coverage (say less than 500x in 90% of the samples). On the basis of this information, we might consider dropping them from the assay. 


```{r filter_data, message=FALSE, warning=FALSE, results='hide'}
# initialize df for flagged samples
df_flagged <- data.frame(matrix(ncol = ncol(bed_coverage), nrow = 0))
colnames(df_flagged) <- colnames(bed_coverage)
# loop through the collated coverage data and check samples coverage for each bed entry
for (region in 1:nrow(bed_coverage)){
  counter <- 0
  for (sample in 6:ncol(bed_coverage)){
    if (bed_coverage[region,sample] < 500.0) {
      counter = counter + 1
    }
  }
  # checking if the region has less than 500x coverage in 90% of the samples
  if (counter/length(list_of_files) > 0.90){
    df_flagged[nrow(df_flagged) + 1,] <- bed_coverage[region,]
  }
}
```

```{r display_data, message=FALSE, warning=FALSE}
# display filtered data in a table
datatable(df_flagged, rownames = FALSE, filter="top", options = list(pageLength = 10, scrollX=T, scrollY=T), caption = "Regions consistently with low coverage <1000x")
write.csv(df_flagged[ ,1:4], file = 'stitched_tst500c_low_coverage_regions.csv', row.names = FALSE, quote = FALSE)
```

Out of `r nrow(bed_coverage)` regions in the target bed file, `r nrow(df_flagged)` regions shave less than 1000x coverage in 90% of the (`r length(list_of_files)`) samples.

## TERT promoter region check

```{r tert, message=FALSE, warning=FALSE, results='hide'}
tert_row <-  which(bed_coverage$name == "TERT_Promoter")
tert_row_info <- as.numeric(bed_coverage[tert_row,6:ncol(bed_coverage)])
length(tert_row_info[tert_row_info>1000])

hist_row <-  which(bed_coverage$name == "HIST2H3A_Exon1_NM001005464")
hist_row_info <- as.numeric(bed_coverage[hist_row,6:ncol(bed_coverage)])
length(hist_row_info[hist_row_info>1000])
```

## Comparison with vcf analysis

```{r compare_vcf, message=FALSE, warning=FALSE}
samples_hm <- scan("/Users/kanwals/UMCCR/data/projects/cttso-qc/samples_zero_cov_helen.txt", what="", sep="\n")
regions_hm <- as.data.frame(read.table("/Users/kanwals/UMCCR/data/projects/cttso-qc/TST500C_manifest_hm.bed", header = FALSE, sep="\t", stringsAsFactors=FALSE))
names.use <- names(bed_coverage)[(names(bed_coverage) %in% samples_hm)]
#subset data to samples of interest and then also subset to regions of interest
bed_coverage_subset <-  bed_coverage[, names.use] %>%
  filter(name %in% regions_hm$V4)

regions_hm %>%
  filter(!regions_hm$V4 %in% bed_coverage$name)
```

