---
title: "Coverage analysis of cttso data"
author: "Sehrish Kanwal"
date: "`r format(Sys.time(), '%a %Y-%b-%d')`"
output:
  html_document:
    keep_md: yes
    css: style.css
    theme: readable
    toc: true
    toc_float: true
    code_folding: hide
  rmdformats::material:
    highlight: kate
params:
  mosdepth_output: '/Users/kanwals/UMCCR/data/projects/cttso-qc/mosdepth_output_tst500c'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
### Required packages.
library(here)
library(dplyr)
library(tidyverse)
library(stringr)
library(DT)
library(jtools)
```

***

## Introduction

In this analysis, we are looking at the coverage of panel target regions across our tso cohort.

## Data

The TSO500 alignment files used for calculating coverage exist in GDS. The script used to run [mosdepth tool](https://github.com/brentp/mosdepth) on these bams for ~280 samples can be found [here](./scripts/run_cttso_bams_through_mosdepth.sh)

mosdepth can output the mean per-region given a BED file of regions. The result is summarised in the table below:


```{r raw_cov_results, message=FALSE, warning=FALSE}
# read in bed file with regions of interest
bed_coverage <- as.data.frame(read.table("/Users/kanwals/UMCCR/data/projects/cttso-qc/TST500C_manifest.bed", header = FALSE, sep="\t", stringsAsFactors=FALSE))
colnames(bed_coverage) <- c("chr", "region_start", "region_end", "name")

# list files in the input directory - made using mosdepth
list_of_files <- list.files(path = params$mosdepth_output,
                            recursive = TRUE,
                            pattern = "\\.bed.gz$",
                            full.names = TRUE)

for (i in 1:length(list_of_files)){
  # read files and prepare data format for display
  file <- read_tsv(list_of_files[i],col_names = c("chr", "region_start", "region_end", "name", "mean_coverage" ),col_types = "cn") %>%
      mutate(sample_id = basename(list_of_files[i])) %>%
      mutate(sample_id = gsub("*.regions.bed.gz", "", sample_id)) 
  # find sample id and rename mean_coverage column using it
  sample_name <- unique(file$sample_id)
  sample_coverage <- file %>%
    dplyr::rename(!!sample_name := mean_coverage) %>%
    dplyr::select(one_of(sample_name))
  # add sample_coverage to original bed file
  bed_coverage <- cbind(bed_coverage, sample_coverage) 
}

```

## Coverage check

The most relevant information would be what regions consistently have less than X coverage (say less than 1000x in 90% of the samples). That would convince me to drop them from the assay. 

```{r filter_data, message=FALSE, warning=FALSE, results='hide'}
# initialize variables for loops
df_flagged <- data.frame(matrix(ncol = ncol(bed_coverage), nrow = 0))
colnames(df_flagged) <- colnames(bed_coverage)
for (region in 1:nrow(bed_coverage)){
  counter <- 1
  for (sample in 5:length(list_of_files)){
    if (bed_coverage[region,sample] < 1000.0) {
      counter = counter + 1
    }
  }
  if (counter/length(list_of_files) > 0.90){
    df_flagged[nrow(df_flagged) + 1,] <- bed_coverage[region,]
  }
}

```

