---
title: "Tumour Mutational Burden"
author: "Sehrish Kanwal"
date: "`r format(Sys.time(), '%a %Y-%b-%d')`"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE) # global setup
```

Required packages

```{r, message=FALSE}
library(vcfR)
library(dplyr)
library(ggplot2)
```

Input file

```{r}
vcf <-  read.vcfR("../Data/ensemble-pon-pass.vcf")
#head(vcf)
```

Have a named vector with chr + chr_length.
You can extract the chr_length for chr, and then bin based on that.
In a named vector, you can access each element by name such as chr_lengths[chr_you_want].

```{r}
chr_lengths <- c(249250621L, 243199373L, 198022430L, 191154276L,
                 180915260L, 171115067L, 159138663L, 146364022L,
                 141213431L, 135534747L, 135006516L, 133851895L,
                 115169878L, 107349540L, 102531392L,  90354753L,
                 81195210L,  78077248L,   59128983L,  63025520L,
                 48129895L,  51304566L,  155270560L,  59373566L, 16569L)
chr_names = c(1:22, "X", "Y", "MT")
names(chr_lengths) <- chr_names
```

Declare bin size (in MB)

```{r}
bin_size <- 40000000
```

Get chromosome and filter vcf for that chromosome

```{r}
filter_chr <- function(x, vcf) {
  vcf_chr_pos <- data.frame(chr=getCHROM(vcf), pos=getPOS(vcf), stringsAsFactors = FALSE)
  return(filter(vcf_chr_pos, chr == x))
}
```

Prepare bins for the specific chromosome

```{r}
bin_chr <- function(chr_lengths, chr_name, bin_size) {
  return(seq(from = 0, to = chr_lengths[chr_name], by = bin_size))
}
```

Count mutations in a chromosome

```{r}
count_mut_in_chr <- function(chr_pos, bin_vec) {
  x <- cbind(chr_pos$pos, findInterval(chr_pos$pos, bin_vec))
  mean(table(x[, 2]))
}
```

Apply same function to all chromosomes. This gives a numerical vector containing the number of variants for each chromosome 

```{r}
results <- vector("numeric", length = length(chr_names))
for (i in 1:length(chr_names)) {
  results[i] <- count_mut_in_chr(filter_chr(chr_names[i], vcf), bin_chr(chr_lengths, chr_names[i], bin_size))
}
print(results)
```

Plotting

```{r}
chrom <- vcfR::getCHROM(vcf)
x <- table(chrom)

# How can I sort not alphanumerically
# Use gtools::mixedorder to get the ordered indices (or mixedsort to sort them)
sorted_ind <- gtools::mixedorder(names(x))
x <- x[sorted_ind]
chr_fac <- factor(names(x), levels = gtools::mixedsort(names(x)))
df <- data.frame(chr = chr_fac, mut_num = c(unname(x)))
ggplot(df, aes(x = chr, y = mut_num)) +
  geom_point(colour = "purple") +
  theme_bw()
```






















