---
title: "Coverage assessment using mosdepth"
author: "Sehrish Kanwal"
date: "`r format(Sys.time(), '%a %Y-%b-%d')`"
output:
  html_document: 
    keep_md: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE) # global setup
```

**Required packages**

```{r, message=FALSE}
library(knitr)
library(dplyr)
library(kableExtra)
```

```{r}
lf <- function(...) {
  data.frame(fname = list.files(...)) %>% 
    knitr::kable(row.names = TRUE)
}
```

**Input data**

As an example of a decent cellaularity sample, I have selected `CCR180136_WH18F001P025 - purity:0.66` (https://trello.com/c/wuChCgKg/209-201624918whp025).

**Mosdepth run command**

Currently, running mosdepth on spartan, using the following command:

```
mosdepth --no-per-base --by $target --quantize 0:10:50:500: $sample $bam

--no-per-base     dont output per-base depth. skipping this output will speed execution
                  substantially. prefer quantized or thresholded values if possible. 
                  
--by              the $target file - currently using the hotspot regions.

--quantize        allows splitting coverage into bins and merging adjacent regions 
                  that fall into the same bin even if they have different exact coverage values.
```

The location of run directory on spartan is `/data/cephfs/punim0010/projects/Kanwal/Kanwal_mosdepth/samplename`.

**Mosdepth output**

Currently focussing on `regions.bed` file that contains information about coverage of hotspot regions and `quantized.bed` file that contains information about callability windows.

1. Hotspot regions

```{r, message=FALSE}
mosdepth <- "/Users/kanwals/Documents/UMCCR/data/mosdepth"
mosdepth_decentcell <- file.path(mosdepth, "CCR180136_WH18F001P025")
#call lf function to list files in the directory
lf(mosdepth_decentcell)

#read regions.bed file - first grep the appropriate bed file from folder and then create a path for that name
regions_bed_name <- list.files(mosdepth_decentcell)[grep(".regions.bed$",list.files(mosdepth_decentcell))]
regions_bed_path <- file.path(mosdepth_decentcell, regions_bed_name)
regions_bed <- as.data.frame(read.table(regions_bed_path, header = FALSE, sep = "\t", stringsAsFactors = FALSE))

#read hotspots file - to include the column that shows the change from ref. nucleaotide in the regions.bed file
hotspots <- as.data.frame(read.table("~/Documents/UMCCR/data/mosdepth/hotspot.edited.bed", header = FALSE, sep = "\t", stringsAsFactors = FALSE))

#merge region_bed file with hotspots column that has info about the possible nucleotide change. Also avoid adding an additional Row.names column by setting it to NULL
regions_bed_hotspot <- transform(merge(regions_bed, hotspots$V5, by = "row.names", sort = FALSE), Row.names=NULL) %>%
  select(V1, V2, V4, y, V5)
#assign column names  
colnames(regions_bed_hotspot) <- c("Chrom","Pos", "Ref", "Alt", "Cov")

#present output in a table format
regions_bed_hotspot %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F) %>%
  scroll_box(height = "400px")
```

2. Callability levels

```{r, message=FALSE}
#read quatized.bed file - first grep the appropriate bed file from folder and then create a path for that name
quantized_bed_name <- list.files(mosdepth_decentcell)[grep(".quantized.bed$",list.files(mosdepth_decentcell))]
quantized_bed_path <- file.path(mosdepth_decentcell, quantized_bed_name)
quantized_bed <- as.data.frame(read.table(quantized_bed_path, header = FALSE, sep = "\t", stringsAsFactors = FALSE))
#assign column names
colnames(quantized_bed) <- c("Chr", "Start", "End", "Callability")

#present output in a table format - not a good idea as takes forever to render becuase of the number of observations
#quantized_bed %>%
#  kable() %>%
#  kable_styling(bootstrap_options = "striped", full_width = F) %>%
#  scroll_box(height = "400px")

#Extract no coverage regions from quantized.bed file
no_coverage <- quantized_bed[quantized_bed$Callability == "0:10", ]
#Extract low coverage regions from quantized.bed file
low_coverage <- quantized_bed[quantized_bed$Callability == "10:50", ]
#Extract high coverage regions from quantized.bed file
high_coverage <- quantized_bed[quantized_bed$Callability == "500:inf", ]
```






